Analyzing difficulty of Wordle using linguistic characteristics to determine

### 2022-04-28 outputs/paper/paper.Rmd
test

---
title: Analyzing difficulty of Wordle using linguistic characteristics to determine
  average success of Twitter players
author: "Ivan Li"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
    pdf_document:
        keep_tex: true
abstract: Using the Twitter API, data was gathered recording the public scores of
  Wordle players over one month. The frequency of each daily score was obtained and
  analyzed based on level of success, geographical location, and linguistic properties
  of the word. It was found that the linguistic properties of words such as commonality
  have impacts on player success. These results may have significance in literary
  and linguistic analysis.
thanks: 'Code and data are available at: https://github.com/Ivannoar/Twilight.'
toc: no
bibliography: references.bib
---

@@

library(tidyverse)
library(kableExtra)
library(countrycode)
library(chron)
library(ds4psy)
library(scales)
library(DiagrammeR)
library(knitr)
library(tinytex)

@@

```{r preview, tab.cap = "Sample of Cleaned Twitter Dataset", tab.width=5, echo=FALSE}
masterdata_clean |>
  select(date_tweeted, country, score, hardmode) |>
  arrange(date_tweeted) |>
  slice(1:10) |>
  kable(
    caption = "First ten rows of a dataset showing Wordle-related tweets",
    col.names = c("Date", "Country", "Wordle Score", "Hardmode"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) |>
  kable_styling(latex_options = c("striped", "scale_down"))

@@

Plots were created corresponding to the chosen variables in our cleaned dataset in order to familiarize with the obtained data and provide motivation for further analysis. To begin with, we wish to plot the frequency of each result/score across our entire dataset. We compare the frequency of each result and plot it as both the raw frequency and the percentage that each score appears in our dataset. Note that failures, where the user was unable to guess the word in the maximum 6 tries, is plotted as a score of 7. We observe that a majority of Twitter users guess complete the game within 4 tries, and that the distribution of guesses is sensibly unimodal. The distribution is slightly right-skewed, and the dataset shows that the number of failures is over three times as much as the number of people who guess the word on their first try, which is expected due to the low probability of guessing the exact word with no prior information compared to the chance of failure. This is also reflected in the proportions between scores of 2 and 6. 
```{r freq, fig.cap = "Scores of Twitter Wordle Players", echo = FALSE}
masterdata_clean |>
  ggplot(mapping = aes(x = score)) + geom_bar(fill = '#6aaa64') + xlab("Score") + ylab("Frequency") +
  geom_text(aes(label = ..count..), vjust = -0.5, stat = "count")

@@

We also show the frequency that countries occur in our dataset, using the geographical data provided in the tweets. We display the top 10 countries found in our dataset and observe that the Americas and the UK make up a majority of the tweets found in the data. 
```{r country, fig.cap = "Top 10 Frequent Wordle Countries", echo = FALSE}
countryfreq <- table(masterdata_clean$country)
countryfreq |>
  as.data.frame() |>
  arrange(desc(Freq)) |>
  slice(1:10) |>
  kable(
    caption = "Top 10 most common countries found in dataset",
    col.names = c("Country", "Frequency"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) |> kable_styling(latex_options = c("striped", "scale_down"))
```
To conduct further analysis, we seek to plot data based on the average score of users on different dates and different countries. Shown is a table displaying the date of the daily word being given, the average score of Twitter users across our dataset, and the word of the day. Also plotted is the distribution of the 'easiest' and 'hardest' words determined by average score, displayed in percentages and grouped by the mode used.
```{r wordinfo, fig.cap = "Score and Word by Date", echo = FALSE}

average_score_df <- masterdata_clean |>
  group_by(date_tweeted) |>
  transmute(average_score = mean(as.numeric(score))) |>
  unique() |>
  arrange(date_tweeted, desc(date_tweeted))
average_score_df <- average_score_df[-c(8, 9, 17, 18), ]

@@

#average_score_df$daily_word |>
  #count_chars()

average_score_df |> 
  kable(
    caption = "Average score of Twitter Wordle players",
    col.names = c("Date", "Average Score", "Daily Word"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) |> 
  kable_styling(latex_options = c("striped", "scale_down"))

@@

average_score_df |> 
  kable(
    caption = "Linguistic Properties of Daily Wordle Words",
    col.names = c("Date", "Average Score", "Daily Word", "Word Frequency Rank", "Orthographic Neighbours"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) |> 
  kable_styling(latex_options = c("striped", "scale_down"))

@@

```{r, echo=FALSE}
masterdata_clean_modelmain <- cbind(masterdata$text, masterdata_clean)|>
  mutate(text = masterdata$text)|>
  subset(select = -c(1))|>
  mutate(text = sub(".*Wordle ", "", text ))|>
  mutate(text = sub(" .*", "", text))

@@

masterdata_clean_modelmain <- masterdata_clean_modelmain|>
  filter(between(text, 289,310) == TRUE)

masterdata_clean_modelmain <- masterdata_clean_modelmain |>
  mutate(rank = ifelse(
    text==289, 11938, ifelse(
    text==290, 26852, ifelse(
    text==291, 11931, ifelse(
    text==292, 10467, ifelse(
    text==293, 3837,  ifelse(
    text==294, 2880,  ifelse(
    text==295, 253,   ifelse(
    text==302, 6374,  ifelse( 
    text==303, 11502, ifelse(
    text==304, 9811,  ifelse(
    text==305, 4953,  ifelse(
    text==306, 10008, ifelse(
    text==307, 623, 6242
 ))))))))))))))

masterdata_clean_modelmain <- masterdata_clean_modelmain |>
  mutate(neighbournum = ifelse(
    text==289, 2, ifelse(
    text==290, 3, ifelse(
    text==291, 1, ifelse(
    text==292, 1, ifelse(
    text==293, 10,ifelse(
    text==294, 2, ifelse(
    text==295, 5, ifelse(
    text==302, 3, ifelse( 
    text==303, 1, ifelse(
    text==304, 2, ifelse(
    text==305, 1, ifelse(
    text==306, 0, ifelse(
    text==307, 5, 1
 ))))))))))))))

masterdata_clean_modelmain_norm <-  masterdata_clean_modelmain|>
  filter(hardmode == "Off")
masterdata_clean_modelmain_hard <-  masterdata_clean_modelmain|>
  filter(hardmode == "On")

@@

masterdata_clean_modelmain_norm <- masterdata_clean_modelmain_norm|>
  mutate(failed = (score == 7))
masterdata_clean_modelmain_hard <- masterdata_clean_modelmain_hard|>
  mutate(failed = (score == 7))

>>>>

---
title: Analyzing difficulty of Wordle using linguistic characteristics to determine
  average success of Twitter players
author: "Ivan Li"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
abstract: Using the Twitter API, data was gathered recording the public scores of
  Wordle players over one month. The frequency of each daily score was obtained and
  analyzed based on level of success, geographical location, and linguistic properties
  of the word. It was found that the linguistic properties of words such as commonality
  have impacts on player success. These results may have significance in literary
  and linguistic analysis.
thanks: 'Code and data are available at: https://github.com/Ivannoar/Twilight.'
toc: no
bibliography: references.bib
always_allow_html: yes
---

@@

library(tidyverse)
library(kableExtra)
library(countrycode)
library(chron)
library(ds4psy)
library(scales)
library(DiagrammeR)
library(knitr)
library(tinytex)
library(haven)

@@

```{r preview, tab.cap = "Sample of Cleaned Twitter Dataset", tab.width=5, echo=FALSE}
masterdata_clean %>%
  select(date_tweeted, country, score, hardmode) %>%
  arrange(date_tweeted) %>%
  slice(1:10) %>%
  kable(
    caption = "First ten rows of a dataset showing Wordle-related tweets",
    col.names = c("Date", "Country", "Wordle Score", "Hardmode"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down"))

@@

Plots were created corresponding to the chosen variables in our cleaned dataset in order to familiarize with the obtained data and provide motivation for further analysis. To begin with, we wish to plot the frequency of each result/score across our entire dataset. We compare the frequency of each result and plot it as both the raw frequency and the percentage that each score appears in our dataset. Note that failures, where the user was unable to guess the word in the maximum 6 tries, is plotted as a score of 7. We observe that a majority of Twitter users guess complete the game within 4 tries, and that the distribution of guesses is sensibly unimodal. The distribution is slightly right-skewed, and the dataset shows that the number of failures is over three times as much as the number of people who guess the word on their first try, which is expected due to the low probability of guessing the exact word with no prior information compared to the chance of failure. This is also reflected in the proportions between scores of 2 and 6. 
```{r freq, fig.cap = "Scores of Twitter Wordle Players", echo = FALSE}
masterdata_clean %>%
  ggplot(mapping = aes(x = score)) + geom_bar(fill = '#6aaa64') + xlab("Score") + ylab("Frequency") +
  geom_text(aes(label = ..count..), vjust = -0.5, stat = "count")

@@

We also show the frequency that countries occur in our dataset, using the geographical data provided in the tweets. We display the top 10 countries found in our dataset and observe that the Americas and the UK make up a majority of the tweets found in the data. 
```{r country, fig.cap = "Top 10 Frequent Wordle Countries", echo = FALSE}
countryfreq <- table(masterdata_clean$country)
countryfreq %>%
  as.data.frame() %>%
  arrange(desc(Freq)) %>%
  slice(1:10) %>%
  kable(
    caption = "Top 10 most common countries found in dataset",
    col.names = c("Country", "Frequency"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) %>% kable_styling(latex_options = c("striped", "scale_down"))
```
To conduct further analysis, we seek to plot data based on the average score of users on different dates and different countries. Shown is a table displaying the date of the daily word being given, the average score of Twitter users across our dataset, and the word of the day. Also plotted is the distribution of the 'easiest' and 'hardest' words determined by average score, displayed in percentages and grouped by the mode used.
```{r wordinfo, fig.cap = "Score and Word by Date", echo = FALSE}

average_score_df <- masterdata_clean %>%
  group_by(date_tweeted) %>%
  transmute(average_score = mean(as.numeric(score))) %>%
  unique() %>%
  arrange(date_tweeted, desc(date_tweeted))
average_score_df <- average_score_df[-c(8, 9, 17, 18), ]

@@

#average_score_df$daily_word %>%
  #count_chars()

average_score_df %>% 
  kable(
    caption = "Average score of Twitter Wordle players",
    col.names = c("Date", "Average Score", "Daily Word"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) %>% 
  kable_styling(latex_options = c("striped", "scale_down"))

@@

average_score_df %>% 
  kable(
    caption = "Linguistic Properties of Daily Wordle Words",
    col.names = c("Date", "Average Score", "Daily Word", "Word Frequency Rank", "Orthographic Neighbours"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) %>% 
  kable_styling(latex_options = c("striped", "scale_down"))

@@

```{r, echo=FALSE, include=FALSE}
masterdata_clean_modelmain <- cbind(masterdata$text, masterdata_clean)%>%
  mutate(text = masterdata$text)%>%
  subset(select = -c(1))%>%
  mutate(text = sub(".*Wordle ", "", text ))%>%
  mutate(text = sub(" .*", "", text))

@@

masterdata_clean_modelmain <- masterdata_clean_modelmain%>%
  filter(between(text, 289,310) == TRUE)

masterdata_clean_modelmain <- masterdata_clean_modelmain %>%
  mutate(rank = ifelse(
    text==289, 11938, ifelse(
    text==290, 26852, ifelse(
    text==291, 11931, ifelse(
    text==292, 10467, ifelse(
    text==293, 3837,  ifelse(
    text==294, 2880,  ifelse(
    text==295, 253,   ifelse(
    text==302, 6374,  ifelse( 
    text==303, 11502, ifelse(
    text==304, 9811,  ifelse(
    text==305, 4953,  ifelse(
    text==306, 10008, ifelse(
    text==307, 623, 6242
 ))))))))))))))

masterdata_clean_modelmain <- masterdata_clean_modelmain %>%
  mutate(neighbournum = ifelse(
    text==289, 2, ifelse(
    text==290, 3, ifelse(
    text==291, 1, ifelse(
    text==292, 1, ifelse(
    text==293, 10,ifelse(
    text==294, 2, ifelse(
    text==295, 5, ifelse(
    text==302, 3, ifelse( 
    text==303, 1, ifelse(
    text==304, 2, ifelse(
    text==305, 1, ifelse(
    text==306, 0, ifelse(
    text==307, 5, 1
 ))))))))))))))

masterdata_clean_modelmain_norm <-  masterdata_clean_modelmain%>%
  filter(hardmode == "Off")
masterdata_clean_modelmain_hard <-  masterdata_clean_modelmain%>%
  filter(hardmode == "On")

@@

masterdata_clean_modelmain_norm <- masterdata_clean_modelmain_norm%>%
  mutate(failed = (score == 7))
masterdata_clean_modelmain_hard <- masterdata_clean_modelmain_hard%>%
  mutate(failed = (score == 7))
